Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations

### 2024-08-19 main.tex
minor updates

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

@@

% DONE \todo{Reformulate the first paragraph. } % 
Potentials for the differential operators of vector calculus and exterior calculus are of fundamental importance.
The operator norms of the potentials are upper bounds for the Poincar\'e--Friedrichs constants, 
which quantify the fundamental stability features of a variety of partial differential equations.
% which serve a variety of purposes:
% these constants quantify the fundamental stability properties of partial differential equations, imply lower bounds for the spectra of vector and Hodge Laplacians, and they . 
However, while potential operators and Poincar\'e--Friedrichs constants for the gradient have been subject to extensive study, quantifiable results regarding the curl and divergence operators or the exterior derivative are not nearly as numerous yet. 
% We study potential operators in the special class of domains with shellable triangulations,
% which include local stars in 2D and 3D. 
% We bound the operator norms of our operators in terms of the shape regularity of the triangulation.

@@

We wish to clarify the relationship between the Poincar\'e--Friedrichs inequality, in the sense introduced above, with other inequalities that are known as Poincar\'e inequality (or also Poincar\'e--Wirtinger or Friedrichs inequality) in the literature. Given $p \in [1,\infty]$ and an open set $\Omega \subseteq \bbR^{n}$ of finite measure, \todo{cite Ern...}

@@

\todo{MWL: I have removed a few lines to shorten the discussion. We should discuss later how to set up this section and what results to present to keep a good flow.}

@@

% DONE: Introduce a reference simplex 
% DONE: DEFINE THE JACOBIAN 
\todo{MWL: If we need to define the Jacobian formally, where should that be put? Maybe an extra paragraph \textbf{Notation} immediately at the end of the introduction section.}

@@

    Suppose that $u \in W^{1,p}(U)$. Then $\hat u := u \circ \varphi \in W^{1,p}(\hat U)$. 
    We observe 
    \begin{align*}
        \| \nabla \hat u \|_{L^{p}(\hat U)}^{p}
        &=
        \| \nabla \hat u \|_{L^{p}(\hat T_1)}^{p}
        +
        \| \nabla \hat u \|_{L^{p}(\hat T_2)}^{p}
        \\&
        \leq 
        |\det(\Jacobian \varphi_1^{-1})|
        \cdot \| \Jacobian \varphi_1 \|^{p}
        \cdot 
        \| \nabla u \|_{L^{p}(T_1)}^{p}
        +
        |\det(\Jacobian \varphi_2^{-1})|
        \cdot \| \Jacobian \varphi_2 \|^{p}
        \cdot 
        \| \nabla u \|_{L^{p}(T_2)}^{p}
        \\&
        \leq 
        \max\left( 
            |\det(\Jacobian \varphi_1)|^{-1} \cdot \| \Jacobian \varphi_1 \|^{p}
            ,
            |\det(\Jacobian \varphi_2)|^{-1} \cdot \| \Jacobian \varphi_2 \|^{p}
        \right)
        \| \nabla u \|_{L^{p}(U)}^{p}
        .
    \end{align*}
    Hence 
    \begin{align*}
        \| \nabla \hat u \|_{L^{p}(\hat U)}
        &\leq 
        \max\left( 
            |\det(\Jacobian \varphi_1)|^{-\frac 1 p} \cdot \| \Jacobian \varphi_1 \|
            ,
            |\det(\Jacobian \varphi_2)|^{-\frac 1 p} \cdot \| \Jacobian \varphi_2 \|
        \right)
        \| \nabla u \|_{L^{p}(U)}
        .
    \end{align*}
    Notice that $\hat U$ has diameter $2$. 
    There exists $\hat w \in W^{1,p}(\hat U)$ such that $\nabla \hat w = \nabla \hat u$ and 
    \begin{align*}
        \| \hat w \|_{L^{p}(\hat U)}
        \leq 
        2 C_{{\rm PF},{\rm EFNT},p}
        \| \nabla \hat u \|_{L^{p}(\hat U)}
        .
    \end{align*}
    Next, setting $w := \hat w \circ \varphi^{-1}$, we find $\nabla w = \nabla u$ and 
    \begin{align*}
        \| w \|_{L^{p}(U)}
        \leq 
        \max\left( 
            |\det(\Jacobian \varphi_1)|
            ,
            |\det(\Jacobian \varphi_2)|
        \right)^{\frac 1 p}
        \| \hat w \|_{L^{p}(\hat U)}
        .
    \end{align*}
    % DONE: that abbreviation is to close to the Jacobian / total derivative 
    Abbreviate $D := \max\left( \diam(T_1), \diam(T_2) \right)$. In combination, 
    \begin{align*}
        \| w \|_{L^{p}(U)}
        &\leq 
        2 C_{{\rm PF},{\rm EFNT},p}
        \max\left( 
            |\det(\Jacobian \varphi_1)|
            ,
            |\det(\Jacobian \varphi_2)|
        \right)^{\frac 1 p}
        \max\left( 
            |\det(\Jacobian \varphi_1)|^{-\frac 1 p} 
            ,
            |\det(\Jacobian \varphi_2)|^{-\frac 1 p} 
        \right)
        D
        \| \nabla u \|_{L^{p}(U)}
        \\&
        \leq 
        2 C_{{\rm PF},{\rm EFNT},p}
        \max\left( 
            |\det(\Jacobian \varphi_1)|
            ,
            |\det(\Jacobian \varphi_2)|
        \right)^{\frac 1 p}
        \min\left( 
            |\det(\Jacobian \varphi_1)|
            ,
            |\det(\Jacobian \varphi_2)| 
        \right)^{-\frac 1 p} 
        D
        \| \nabla u \|_{L^{p}(U)}
        \\&
        \leq 
        2 C_{{\rm PF},{\rm EFNT},p}
        \max\left( 
            \frac{ |\det(\Jacobian \varphi_1)| }{ |\det(\Jacobian \varphi_2)| }
            ,
            \frac{ |\det(\Jacobian \varphi_2)| }{ |\det(\Jacobian \varphi_1)| }
        \right)^{\frac 1 p}
        D
        \| \nabla u \|_{L^{p}(U)}
        \\&
        \leq 
        2 C_{{\rm PF},{\rm EFNT},p}
        \max\left( 
            \frac{ \vol(T_1) }{ \vol(T_2) }
            ,
            \frac{ \vol(T_2) }{ \vol(T_1) }
        \right)^{\frac 1 p}
        D
        \| \nabla u \|_{L^{p}(U)}
        .
    \end{align*}
    This gives the desired result. 
\end{proof}

@@

%     \node at (A) [below left] {$A$};
%     \node at (B) [below right] {$B$};
%     \node at (C) [above] {$C$};
%     \node at (D) [above right] {$D$};
%     \node at (T) [above right] {$T$};
\end{tikzpicture}
\end{center}
\caption{Strongly connected triangulation of a domain. The arrows depict a tree in the strong connection graph of the form used in Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.}
\end{figure}

\begin{theorem}\label{theorem:poincarefriedrichsestimate:grad}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a strongly-connected finite triangulation $\calT$.
    Let $T_0 \in \calT$ be an $n$-simplex such that for all $n$-simplices $T \in \calT$ 
    we fix a strong path $T_0, T_1, \dots, T_m = T$.
    Suppose $1 \leq p,q \leq \infty$ with $1 = 1/p + 1/q$.
    Then for any $u \in W^{1,p}(\Omega)$ 
    there exists $w \in W^{1,p}(\Omega)$ with $\nabla u = \nabla w$ 
    and such that for all $T \in \calT$:
    \todo{Make sure the notation for the PF constants is uniform throughout the paper}
    \begin{align*} 
        % \color{red}
        \| w \|_{L^{p}(T_m)}
        &\leq 
        \left( 1 + \rho(\calT)^{\frac q p} \right)^{\frac 1 q}
        \sum_{t=1}^{m}
        \rho(\calT)^{\frac{m-t}{p}}
        C_{{\rm{PF}},T_{t} \cup T_{t-1},p}
        % 2 C_{{\rm PF},{\rm EFNT},p} \rho(\calT)^{\frac 1 p} \max\left( \diam(T_{t}), \diam(T_{t-1}) \right)
        \| \nabla u \|_{L^{p}(T_{t} \cup T_{t-1})}
        \\&\quad \quad \quad 
        +
        \rho(\calT)^{\frac{m-1}{p}}
        % C_{{\rm PF},{\rm EFNT},p} \diam(T_0)
        C_{{\rm{PF}},T_{0},p}
        \| \nabla u \|_{L^{p}(T_{0})}
        .
    \end{align*}
\end{theorem}

@@

 \todo{MLW, MV: add a figure here to illustrate how the domain is exhausted.}

@@

 We assess the norm of this potential. We observe 
 \begin{align*}
    \| w'_{F_m} \|_{L^{p}(T_m)}
    \leq 
    \| u_{F_m} \|_{L^{p}(T_m)}
    +
    \| c_{m} \|_{L^{p}(T_m)},
    \\
    \| c_{m} \|_{L^{p}(T_m)}
    = 
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
    \| c_{m} \|_{L^{p}(T_{{m-1}})},
    \\ 
    \| c_{m} \|_{L^{p}(T_{{m-1}})}
    \leq 
    \| w_{m-1}\|_{L^{p}(T_{{m-1}})} + \| u_{F_{m}} \|_{L^{p}(T_{{m-1}})} 
    .
 \end{align*}
 In particular,
 \begin{align*}
    \| w_{m} \|_{L^{p}(T_m)}
    \leq 
    \| u_{F_m} \|_{L^{p}(T_m)}
    +
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
    \| u_{F_{m}} \|_{L^{p}(T_{{m-1}})}
    +
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
    \| w_{m-1}\|_{L^{p}(T_{{m-1}})}
    .
 \end{align*}
 We sum the two integrals of $u_{F_m}$. 
 % 
 In the special case $p=1$,
 \begin{align*}
    \| w_{m} \|_{L^{1}(T_m)}
    \leq 
    \max\left(
        1, \frac{ \vol(T_m) }{ \vol(T_{{m-1}}) } 
    \right)
    \| u_{F_m} \|_{L^{1}(U_{F_m})}
    +
    \frac{ \vol(T_m) }{ \vol(T_{{m-1}}) }
    \| w_{m-1}\|_{L^{1}(T_{{m-1}})}
    .
 \end{align*}
 In the special case $p=\infty$, 
 \begin{align*}
    \| w_{m} \|_{L^{\infty}(T_m)}
    \leq 
    2
    \| u_{F_m} \|_{L^{\infty}(U_{F_m})}
    +
    \| w_{m-1}\|_{L^{\infty}(T_{{m-1}})}
    .
 \end{align*}
 % 1/q = 1 - 1/p = (p-1)/p 
 When $1 < p < \infty$, letting $q := p/(p-1) \in [1,\infty]$ be the complementary exponent, we observe 
 \begin{align*}
    \| w_{m} \|_{L^{p}(T_m)}
    &\leq 
    \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{m-1}})^{\frac q p} } \right)^{\frac 1 q}
    \left( 
        \| u_{F_m} \|_{L^{p}(T_m)}^{p}
        +
        \| u_{F_{m}} \|_{L^{p}(T_{{m-1}})}^{p}
    \right)^{\frac 1 p}
    +
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
    \| w_{m-1}\|_{L^{p}(T_{{m-1}})}
    \\&\leq 
    \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{m-1}})^{\frac q p} } \right)^{\frac 1 q}
    \| u_{F_m} \|_{L^{p}(U_{F_m})} 
    +
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
    \| w_{m-1}\|_{L^{p}(T_{{m-1}})}
    .
 \end{align*}
 Recursive application provides the following estimate:
%  if $T_0 = S_0, S_1, \dots, S_m = T_m$ is a sequence of $n$-simplices in $\calT$
%  such that \color{red}$S_{t-1}$ precedes $S_{t}$ in the graph \color{black} and such that $F_t = S_t \cap S_{t-1}$,
%  then 
 \begin{align*}
    \| w_{m} \|_{L^{p}(T_m)}
    &\leq 
    \sum_{t=1}^{m} 
    \left( 
        \prod_{t < l \leq m} 
        \frac{ \vol(T_l)^{\frac 1 p} }{ \vol(T_{l-1})^{\frac 1 p} } 
    \right)
    \left( 1 + \frac{ \vol(T_t)^{\frac q p} }{ \vol(T_{t-1})^{\frac q p} } \right)^{\frac 1 q}
    \| u_{F_t} \|_{L^{p}(U_{F_t})}
    \\&\quad \quad \quad 
    +
    \left( 
        \prod_{0 < l \leq m} 
        \frac{ \vol(T_l)^{\frac 1 p} }{ \vol(T_{l-1})^{\frac 1 p} } 
    \right)
    \| u_{T_0}\|_{L^{p}(T_{0})}
    .
 \end{align*}
 Taking into account Definition~\eqref{math:moremeasures}, we obtain the desired inequality.
%  \begin{align*}
%     \| w_{m} \|_{L^{p}(T_m)}^{p}
%     &\leq 
%     (m+1)^{p-1}
%     % \sum_{t=1}^{m} \rho(\calT)^{m-t} 
%     % \sum_{t=0}^{m-1} \rho(\calT)^{t} 
%     \frac{\rho(\calT)^{m}-1}{\rho(\calT)-1}
%     \left( 1 + \rho(\calT)^{\frac q p} \right)^{p/q}
%     \| u_{F_t} \|_{L^{p}(U_{F_t})}^{p}
%     \\&\quad \quad \quad 
%     +
%     (m+1)^{p-1}
%     \rho(\calT)^{m-1}
%     \| u_{T_0}\|_{L^{p}(T_{0})}^{p}
%     .
%  \end{align*}
%  Notice that $p/q = p - 1$.
 % 
 Repeating this argument produces $w \in L^{p}(\Omega)$ 
 such that for all $n$-simplices $T \in \calT$ we have the restriction $w_{|T} \in W^{1,p}(T)$ with $w_{|T} = u + c_{0}$.
 In particular, $w \in W^{1,p}(\Omega)$ with $\nabla w = \nabla u$.
\end{proof}

@@

\begin{lemma}    
    Suppose that an $n$-dimensional simplicial complex $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$ and is non-branching, 
    that is, each simplex of dimension $n-1$ is contained in at most two simplices. 
    % DONE: what is non-branching?
    Then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Suppose that $X_m := T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a topological $n$-ball and that $T_{m+1}$ is the next $n$-simplex in the shelling.
    By definition, $I_{m+1} := X_{m} \cap T_{m+1}$ is a submanifold of $\partial T_{m+1}$,
    and it is triangulated by some faces of $T_{m+1}$ and their subsimplices.

@@

\begin{example}
    Any simplex $T$ trivially has a shelling. The boundary complex $\partial \calT(T)$ has a shelling:
    in fact, any enumeration of the boundary faces of $T$ constitutes a shelling.
\end{example}
\begin{example}
    The standard triangulation of the $n$-dimensional cube by six tetrahedra, the Kuhn triangulation~\cite{kuhn1960some}, is shellable.~\footnote{We remark that Kuhn attributes this triangulation to Lefschetz~\cite{lefschetz2015introduction}.}
\end{example}

@@

\todo{cite Guzman and Salgado at the right spot.}\\

>>>>

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
\newcommand{\notice}[1]{\textcolor{red}{REMARK: #1}}

@@

% DONE {Reformulate the first paragraph. } % 
Potentials for the differential operators of vector calculus and exterior calculus are of fundamental importance.
The operator norms of the potentials are upper bounds for the Poincar\'e--Friedrichs constants, 
which quantify the fundamental stability features of a variety of partial differential equations.
% which serve a variety of purposes:
% these constants quantify the fundamental stability properties of partial differential equations, imply lower bounds for the spectra of vector and Hodge Laplacians, and they . 
However, while potential operators and Poincar\'e--Friedrichs constants for the gradient have been subject to extensive study, quantifiable results regarding the curl and divergence operators or the exterior derivative are not nearly as numerous yet. 
% We study potential operators in the special class of domains with shellable triangulations,
% which include local stars in 2D and 3D. 
% We bound the operator norms of our operators in terms of the shape regularity of the triangulation.

@@

We wish to clarify the relationship between the Poincar\'e--Friedrichs inequality, in the sense introduced above, with other inequalities that are known as Poincar\'e inequality (or also Poincar\'e--Wirtinger or Friedrichs inequality) in the literature~\cite{ern2021finite}. Given $p \in [1,\infty]$ and an open set $\Omega \subseteq \bbR^{n}$ of finite measure, 

@@

% \notice{MWL: I have removed a few lines to shorten the discussion. We should discuss later how to set up this section and what results to present to keep a good flow.}

@@

% DONE: Introduce a reference simplex 
% DONE: DEFINE THE JACOBIAN 
% \notice{MWL: If we need to define the Jacobian formally, where should that be put? Maybe an extra paragraph \textbf{Notation} immediately at the end of the introduction section.}

@@

    Suppose that $u \in W^{1,p}(U)$. Then $\hat u := u \circ \varphi \in W^{1,p}(\hat U)$. 
    We observe 
    \begin{align*}
        \| \nabla \hat u \|_{L^{p}(\hat U)}^{p}
        &=
        \| \nabla \hat u \|_{L^{p}(\hat T_1)}^{p}
        +
        \| \nabla \hat u \|_{L^{p}(\hat T_2)}^{p}
        ,
        \\
        \| \nabla \hat u \|_{L^{p}(\hat T_1)}^{p}
        &\leq 
        |\det(\Jacobian \varphi_1^{-1})|
        \cdot \| \Jacobian \varphi_1 \|^{p}
        \cdot 
        \| \nabla u \|_{L^{p}(T_1)}^{p}
        ,
        \\
        \| \nabla \hat u \|_{L^{p}(\hat T_2)}^{p}
        &\leq 
        |\det(\Jacobian \varphi_2^{-1})|
        \cdot \| \Jacobian \varphi_2 \|^{p}
        \cdot 
        \| \nabla u \|_{L^{p}(T_2)}^{p}
%         \\&
%         \leq 
%         \max\left( 
%             |\det(\Jacobian \varphi_1)|^{-1} \cdot \| \Jacobian \varphi_1 \|^{p}
%             ,
%             |\det(\Jacobian \varphi_2)|^{-1} \cdot \| \Jacobian \varphi_2 \|^{p}
%         \right)
%         \| \nabla u \|_{L^{p}(U)}^{p}
        .
    \end{align*}
    Hence 
    \begin{align*}
        \| \nabla \hat u \|_{L^{p}(\hat U)}
        &\leq 
        \max\left( 
            |\det(\Jacobian \varphi_1)|^{-\frac 1 p} \| \Jacobian \varphi_1 \|
            ,
            |\det(\Jacobian \varphi_2)|^{-\frac 1 p} \| \Jacobian \varphi_2 \|
        \right)
        \| \nabla u \|_{L^{p}(U)}
        .
    \end{align*}
    Notice that $\hat U$ has diameter $2$. 
    There exists $\hat w \in W^{1,p}(\hat U)$ such that $\nabla \hat w = \nabla \hat u$ and 
    \begin{align*}
        \| \hat w \|_{L^{p}(\hat U)}
        \leq 
        2 C_{{\rm PF},{\rm EFNT},p}
        \| \nabla \hat u \|_{L^{p}(\hat U)}
        .
    \end{align*}
    Next, setting $w := \hat w \circ \varphi^{-1}$, we find $\nabla w = \nabla u$ and 
    \begin{align*}
        \| w \|_{L^{p}(U)}
        \leq 
        \max\left( 
            |\det(\Jacobian \varphi_1)|
            ,
            |\det(\Jacobian \varphi_2)|
        \right)^{\frac 1 p}
        \| \hat w \|_{L^{p}(\hat U)}
        .
    \end{align*}
    % DONE: that abbreviation is to close to the Jacobian / total derivative 
    Abbreviate $D := \max\left( \diam(T_1), \diam(T_2) \right)$. In combination, 
    \begin{align*}
        \| w \|_{L^{p}(U)}
        &\leq 
        2 C_{{\rm PF},{\rm EFNT},p}
        \max\left( 
            |\det(\Jacobian \varphi_1)|
            ,
            |\det(\Jacobian \varphi_2)|
        \right)^{\frac 1 p}
        \\&\qquad 
        \max\left( 
            |\det(\Jacobian \varphi_1)|^{-\frac 1 p} 
            ,
            |\det(\Jacobian \varphi_2)|^{-\frac 1 p} 
        \right)
        D
        \| \nabla u \|_{L^{p}(U)}
        \\&
        \leq 
        2 C_{{\rm PF},{\rm EFNT},p}
        \max\left( 
            |\det(\Jacobian \varphi_1)|
            ,
            |\det(\Jacobian \varphi_2)|
        \right)^{\frac 1 p}
        \\&\qquad 
        \min\left( 
            |\det(\Jacobian \varphi_1)|
            ,
            |\det(\Jacobian \varphi_2)| 
        \right)^{-\frac 1 p} 
        D
        \| \nabla u \|_{L^{p}(U)}
        \\&
        \leq 
        2 C_{{\rm PF},{\rm EFNT},p}
        \max\left( 
            \frac{ |\det(\Jacobian \varphi_1)| }{ |\det(\Jacobian \varphi_2)| }
            ,
            \frac{ |\det(\Jacobian \varphi_2)| }{ |\det(\Jacobian \varphi_1)| }
        \right)^{\frac 1 p}
        D
        \| \nabla u \|_{L^{p}(U)}
        \\&
        \leq 
        2 C_{{\rm PF},{\rm EFNT},p}
        \max\left( 
            \frac{ \vol(T_1) }{ \vol(T_2) }
            ,
            \frac{ \vol(T_2) }{ \vol(T_1) }
        \right)^{\frac 1 p}
        D
        \| \nabla u \|_{L^{p}(U)}
        .
    \end{align*}
    This gives the desired result. 
\end{proof}

@@

%     \node at (A) [below left] {$A$};
%     \node at (B) [below right] {$B$};
%     \node at (C) [above] {$C$};
%     \node at (D) [above right] {$D$};
%     \node at (T) [above right] {$T$};
\end{tikzpicture}
\end{center}
\caption{Strongly connected triangulation of a domain. The arrows depict a tree in the strong connection graph of the form used in Theorem~\ref{theorem:poincarefriedrichsestimate:grad}. The triangle in dark gray is the triangle being added in the current iteration, and the light gray triangles constitute the subdomain assembled so far.}
\end{figure}

\begin{figure}[t]
\begin{center}
\begin{tikzpicture}
    \coordinate (A) at (0,0);   \coordinate (B) at (2,0); \coordinate (C) at (1,1.5);
    \coordinate (D) at (0,2.9); \coordinate (E) at (2.5,3); \coordinate (F) at (3,1.25); 
    \draw[fill=gray!50] (A) -- (B) -- (C) -- cycle; 
    \draw[fill=white  ] (A) -- (C) -- (D) -- cycle; 
    \draw[fill=white  ] (C) -- (D) -- (E) -- cycle; 
    \draw[fill=white  ] (B) -- (C) -- (F) -- cycle; 
\end{tikzpicture}
\quad
\begin{tikzpicture}
    \coordinate (A) at (0,0);   \coordinate (B) at (2,0); \coordinate (C) at (1,1.5);
    \coordinate (D) at (0,2.9); \coordinate (E) at (2.5,3); \coordinate (F) at (3,1.25); 
    \draw[fill=gray!25] (A) -- (B) -- (C) -- cycle; 
    \draw[fill=gray!50] (A) -- (C) -- (D) -- cycle; 
    \draw[fill=white  ] (C) -- (D) -- (E) -- cycle; 
    \draw[fill=white  ] (B) -- (C) -- (F) -- cycle; 
\end{tikzpicture}
\quad
\begin{tikzpicture}
    \coordinate (A) at (0,0);   \coordinate (B) at (2,0); \coordinate (C) at (1,1.5);
    \coordinate (D) at (0,2.9); \coordinate (E) at (2.5,3); \coordinate (F) at (3,1.25); 
    \draw[fill=gray!25] (A) -- (B) -- (C) -- cycle; 
    \draw[fill=gray!25] (A) -- (C) -- (D) -- cycle; 
    \draw[fill=gray!50] (C) -- (D) -- (E) -- cycle; 
    \draw[fill=white  ] (B) -- (C) -- (F) -- cycle; 
\end{tikzpicture}
\quad
\begin{tikzpicture}
    \coordinate (A) at (0,0);   \coordinate (B) at (2,0); \coordinate (C) at (1,1.5);
    \coordinate (D) at (0,2.9); \coordinate (E) at (2.5,3); \coordinate (F) at (3,1.25); 
    \draw[fill=gray!25] (A) -- (B) -- (C) -- cycle; 
    \draw[fill=gray!25] (A) -- (C) -- (D) -- cycle; 
    \draw[fill=gray!25] (C) -- (D) -- (E) -- cycle; 
    \draw[fill=gray!50] (B) -- (C) -- (F) -- cycle; 
\end{tikzpicture}
\end{center}
\caption{Exhausting a domain step by step as in Theorem~\ref{theorem:poincarefriedrichsestimate:grad} and \ref{theorem:poincarefriedrichsestimate:gradzwei}.}
\label{figure:exhausting}
\end{figure}

\begin{theorem}\label{theorem:poincarefriedrichsestimate:grad}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a strongly-connected finite triangulation $\calT$.
    Let $T_0 \in \calT$ be an $n$-simplex such that for all $n$-simplices $T \in \calT$ 
    we fix a strong path $T_0, T_1, \dots, T_m = T$.
    Suppose $1 \leq p,q \leq \infty$ with $1 = 1/p + 1/q$.
    Then for any $u \in W^{1,p}(\Omega)$ 
    there exists $w \in W^{1,p}(\Omega)$ with $\nabla u = \nabla w$ 
    and such that for all $T \in \calT$:
%     DONE {Make sure the notation for the PF constants is uniform throughout the paper}
    \begin{align*} 
        % \color{red}
        \| w \|_{L^{p}(T_m)}
        &\leq 
        \left( 1 + \rho(\calT)^{\frac q p} \right)^{\frac 1 q}
        \sum_{t=1}^{m}
        \rho(\calT)^{\frac{m-t}{p}}
        C_{{\rm{PF}},T_{t} \cup T_{t-1},p}
        % 2 C_{{\rm PF},{\rm EFNT},p} \rho(\calT)^{\frac 1 p} \max\left( \diam(T_{t}), \diam(T_{t-1}) \right)
        \| \nabla u \|_{L^{p}(T_{t} \cup T_{t-1})}
        \\&\quad \quad \quad 
        +
        \rho(\calT)^{\frac{m-1}{p}}
        % C_{{\rm PF},{\rm EFNT},p} \diam(T_0)
        C_{{\rm{PF}},T_{0},p}
        \| \nabla u \|_{L^{p}(T_{0})}
        .
    \end{align*}
\end{theorem}

@@

 % DONE \notice{MWL, MV: add a figure here to illustrate how the domain is exhausted.}

@@

 We assess the norm of this potential. We observe 
 \begin{align*}
    \| w'_{F_m} \|_{L^{p}(T_m)}
    \leq 
    \| u_{F_m} \|_{L^{p}(T_m)}
    +
    \| c_{m} \|_{L^{p}(T_m)},
    \\
    \| c_{m} \|_{L^{p}(T_m)}
    = 
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
    \| c_{m} \|_{L^{p}(T_{{m-1}})},
    \\ 
    \| c_{m} \|_{L^{p}(T_{{m-1}})}
    \leq 
    \| w_{m-1}\|_{L^{p}(T_{{m-1}})} + \| u_{F_{m}} \|_{L^{p}(T_{{m-1}})} 
    .
 \end{align*}
 In particular,
 \begin{align*}
    \| w_{m} \|_{L^{p}(T_m)}
    \leq 
    \| u_{F_m} \|_{L^{p}(T_m)}
    +
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
    \| u_{F_{m}} \|_{L^{p}(T_{{m-1}})}
    +
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
    \| w_{m-1}\|_{L^{p}(T_{{m-1}})}
    .
 \end{align*}
 We sum the two integrals of $u_{F_m}$. 
 % 
 In the special case $p=1$,
 \begin{align*}
    \| w_{m} \|_{L^{1}(T_m)}
    \leq 
    \max\left(
        1, \frac{ \vol(T_m) }{ \vol(T_{{m-1}}) } 
    \right)
    \| u_{F_m} \|_{L^{1}(U_{F_m})}
    +
    \frac{ \vol(T_m) }{ \vol(T_{{m-1}}) }
    \| w_{m-1}\|_{L^{1}(T_{{m-1}})}
    .
 \end{align*}
 In the special case $p=\infty$, 
 \begin{align*}
    \| w_{m} \|_{L^{\infty}(T_m)}
    \leq 
    2
    \| u_{F_m} \|_{L^{\infty}(U_{F_m})}
    +
    \| w_{m-1}\|_{L^{\infty}(T_{{m-1}})}
    .
 \end{align*}
 % 1/q = 1 - 1/p = (p-1)/p 
 When $1 < p < \infty$, letting $q := p/(p-1) \in [1,\infty]$ be the complementary exponent, we observe 
 \begin{align*}
    \| w_{m} \|_{L^{p}(T_m)}
    &\leq 
    \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{m-1}})^{\frac q p} } \right)^{\frac 1 q}
    \left( 
        \| u_{F_m} \|_{L^{p}(T_m)}^{p}
        +
        \| u_{F_{m}} \|_{L^{p}(T_{{m-1}})}^{p}
    \right)^{\frac 1 p}
    \\&\qquad 
    +
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
    \| w_{m-1}\|_{L^{p}(T_{{m-1}})}
    \\&\leq 
    \left( 1 + \frac{ \vol(T_m)^{\frac q p} }{ \vol(T_{{m-1}})^{\frac q p} } \right)^{\frac 1 q}
    \| u_{F_m} \|_{L^{p}(U_{F_m})} 
    +
    \frac{ \vol(T_m)^{\frac 1 p} }{ \vol(T_{{m-1}})^{\frac 1 p} }
    \| w_{m-1}\|_{L^{p}(T_{{m-1}})}
    .
 \end{align*}
 Recursive application provides the following estimate:
%  if $T_0 = S_0, S_1, \dots, S_m = T_m$ is a sequence of $n$-simplices in $\calT$
%  such that \color{red}$S_{t-1}$ precedes $S_{t}$ in the graph \color{black} and such that $F_t = S_t \cap S_{t-1}$,
%  then 
 \begin{align*}
    \| w_{m} \|_{L^{p}(T_m)}
    &\leq 
    \sum_{t=1}^{m} 
    \left( 
        \prod_{t < l \leq m} 
        \frac{ \vol(T_l)^{\frac 1 p} }{ \vol(T_{l-1})^{\frac 1 p} } 
    \right)
    \left( 1 + \frac{ \vol(T_t)^{\frac q p} }{ \vol(T_{t-1})^{\frac q p} } \right)^{\frac 1 q}
    \| u_{F_t} \|_{L^{p}(U_{F_t})}
    \\&\quad \quad \quad 
    +
    \left( 
        \prod_{0 < l \leq m} 
        \frac{ \vol(T_l)^{\frac 1 p} }{ \vol(T_{l-1})^{\frac 1 p} } 
    \right)
    \| u_{T_0}\|_{L^{p}(T_{0})}
    .
 \end{align*}
 Taking into account Definition~\eqref{math:moremeasures}, we obtain the desired inequality.
%  \begin{align*}
%     \| w_{m} \|_{L^{p}(T_m)}^{p}
%     &\leq 
%     (m+1)^{p-1}
%     % \sum_{t=1}^{m} \rho(\calT)^{m-t} 
%     % \sum_{t=0}^{m-1} \rho(\calT)^{t} 
%     \frac{\rho(\calT)^{m}-1}{\rho(\calT)-1}
%     \left( 1 + \rho(\calT)^{\frac q p} \right)^{p/q}
%     \| u_{F_t} \|_{L^{p}(U_{F_t})}^{p}
%     \\&\quad \quad \quad 
%     +
%     (m+1)^{p-1}
%     \rho(\calT)^{m-1}
%     \| u_{T_0}\|_{L^{p}(T_{0})}^{p}
%     .
%  \end{align*}
%  Notice that $p/q = p - 1$.
 % 
 Repeating this argument produces $w \in L^{p}(\Omega)$ 
 such that for all $n$-simplices $T \in \calT$ we have the restriction $w_{|T} \in W^{1,p}(T)$ with $w_{|T} = u + c_{0}$.
 In particular, $w \in W^{1,p}(\Omega)$ with $\nabla w = \nabla u$.
\end{proof}

@@

\begin{lemma}    
    Let $\calT$ be an $n$-dimensional simplicial complex 
    with a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$,
    such that each simplex of dimension $n-1$ is contained in at most two simplices. 
    % DONE: what is non-branching?
    Then
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m}
    $$ 
    is a manifold with boundary for all $0 \leq m \leq M$.
    % 
    In particular, $X_{m}$ is a topological $n$-ball when $m < M$, 
    and 
    $X_{M}$ is either a topological $n$-ball or topological $n$-sphere. 
\end{lemma}
\begin{proof}  
    We prove this claim by induction. 
    Certainly, if $\calT$ contains only one single $n$-simplex, then it is a shellable triangulation of a topological $n$-ball. 
    Suppose that $X_m := T_{0} \cup T_{1} \cup \dots \cup T_{m}$ is a topological $n$-ball and that $T_{m+1}$ is the next $n$-simplex in the shelling.
    By definition, $I_{m+1} := X_{m} \cap T_{m+1}$ is a submanifold of $\partial T_{m+1}$,
    and it is triangulated by some faces of $T_{m+1}$ and their subsimplices.

@@

\begin{example}
    Any simplex $T$ trivially has a shelling. The boundary complex $\partial \calT(T)$ has a shelling:
    in fact, any enumeration of the boundary faces of $T$ constitutes a shelling.
\end{example}
\begin{example}
    The standard triangulation of the $n$-dimensional cube by six tetrahedra, the Kuhn triangulation~\cite{kuhn1960some}, is shellable.\footnote{We remark that Kuhn attributes this triangulation to Lefschetz~\cite{lefschetz2015introduction}.}
\end{example}

@@
